---
title: "Data Mining HW 1"
author: "Veronica Hsieh"
date: "1/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
```
# read orders
```{r}
dat = read.csv("data/orders.csv")
dat$orddt = as.Date(dat$orddate, "%d%b%Y")
head(dat)
dat$orddate = NULL
str(dat)
dim(dat)
```

```{r}
which(is.na(dat))
```

# read dependent variable
```{r}
y = read.csv("data/booktrain.csv")
head(y)
```

# Creating summary variables
# Frequency_ord and monetary_tot look to be highly correlated 
```{r}
x = dat %>%
  group_by(id) %>%
  summarise(
    recency = as.numeric(as.Date('2014-08-01') - max(orddt)),
    frequency = sum(qty),
    frequency_ord = n_distinct(ordnum),
    monetary_tot = sum(price),
    monetary_avg = mean(price),
    tof = as.numeric(as.Date('2014-08-01') - min(orddt))
    ) %>%
  select(id, recency, frequency, frequency_ord, monetary_tot, monetary_avg, tof)
head(x)
cor(x[-1])
```

# merge
```{r}
all = left_join(x, y, by="id")
head(all)
```
# EDA

# histogram for categories
```{r}
hist(dat$category)
```

# plotting summary variables vs logtarg
```{r}
library(ggplot2)
```

```{r}
ggplot(data = all, aes(recency, logtarg)) + 
  geom_point()
ggplot(data = all, aes(frequency, logtarg)) + 
  geom_point()
ggplot(data = all, aes(frequency_ord, logtarg)) + 
  geom_point()
ggplot(data = all, aes(monetary_tot, logtarg)) + 
  geom_point()
ggplot(data = all, aes(monetary_avg, logtarg)) + 
  geom_point()
ggplot(data = all, aes(tof, logtarg)) + 
  geom_point()
```

```{r}
all2 <- all
all2$targ = exp(all$logtarg)-1
```

# plotting summary variables against targ...note most variables are right skewed so may want to take log transformation of variables
```{r}
ggplot(data = all2, aes(recency, targ)) + 
  geom_point()
ggplot(data = all2, aes(frequency, targ)) + 
  geom_point()
ggplot(data = all2, aes(frequency_ord, targ)) + 
  geom_point()
ggplot(data = all2, aes(monetary_tot, targ)) + 
  geom_point()
ggplot(data = all2, aes(monetary_avg, targ)) + 
  geom_point()
ggplot(data = all2, aes(tof, logtarg)) + 
  geom_point()
```

```{r}
fit = lm(logtarg ~ log(frequency + 1), all)
summary(fit)
```

```{r}
fit_spend = lm(targ ~ log(frequency + 1), all2)
summary(fit_spend)
```

```{r}
library(leaps)
full = lm(targ ~ log(recency + 1) + log(tof + 1) + log(frequency + 1) + log(frequency_ord + 1) + log(monetary_tot + 1) + log(monetary_avg + 1),data = all2)
step(full, data = all2, direction = "backward")
```
```{r}
fit_backstep = lm(formula = targ ~ log(recency + 1) + log(tof + 1) + log(frequency + 
    1) + log(frequency_ord + 1) + log(monetary_avg + 1), data = all2)
summary(fit_backstep)
```

# predict spend
```{r}
all2$yhat = predict(fit, all)
```

#transform prediction back to logtarg
```{r}
all$yhat = log(all$yhat_spend + 1)
```

# write results to CSV
```{r}
head(all)
test = is.na(all$logtarg)
write.csv(all[test, c('id', 'yhat')], "output/test_rfm.csv", row.names=F)
```


